Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-11-08 paper/article.tex
Paper progress

The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
consists of a ``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual
features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
the same objects which the monkeys grasped. The outputs are muscle length velocities for the
should, arm, and hand of the monkey during the trial. The natural velocities were captured
with a motion capturing glove, and the artificial neural network is trained to recapitulate
those grasping motions. Data and trained models from this work were supplied to us by the lead
author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.

@@

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
refers to the fact that the activity of the module receiving the visual inputs resembles
the AIP activity of the monkey subject from the same trials. Likewise, the second and
third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
The emergent network dynamics show a number of other correspondences to the monkeys' natural
brain activity as well, as detailed in the paper.

Importantly, the simulated circuit's activity shows a relatively clear separation of the
object shapes. That is - the visual information input to the network differentiates one trial
from another, and the circuit is able to leverage that information to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and
location. The visual information forward-propagates through the network, through successive
processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
for the object represented by the input visual features. It follows, then, that the classes
are separable in some way by observing the distinct ways that the mRNN is activated by the
classes' corresponding visual inputs. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

As a result, simulating a brain lesion in terms of this artificial network results in error
modes which resemble a natural lesion of certain regions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate visual object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. Likewise, if we simulate a
``stroke'' of a portion of the output module of the Michaels mRNN, roughly
corresponding to M1, we see a more wholesale loss of movement, affecting even the
ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.46\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
	    \caption{inner1}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.46\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
	    \caption{inner1}
	\end{subfigure}
	\hfill
	\caption{Two simple graphs}
	\label{fig:lesion}
\end{figure}

* Long-running dynamics

** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
** Cite Michaels, and those upstream from him (Susillo?) re: RNNs. Networks have internal dynamics that result from both inputs and internal connectivity. Perturbations at t=0 affect outputs at t=N.
** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
  simulation to develop and evaluate a neural controller. Use of lesion model.

Example: Use of actual stimulation of an actual brain, together with offline deep learning,
to search for an optimal visual input: \cite{walker.inception}

>>>>

An example of such an approach is Bernal et al. \cite{bernal.sim}. In this work, the
authors use a simulated spiking neural network to train a stimulation agent, much
like the work we present here. Their stimulation agent sought to restore the network's
control of a simulated arm, to reach a target. As they point out, we have a limited
ability to probe a neural circuit \textit{in vivo} in order to perform learning. As
a result, we first need to design our approach through the use of an admissible
simulation. The authors in this case simulated lesions by effectively removing parts
of their simulated network, or by cutting connections between parts of the network.
Likewise here, we use an established design for an artificial neural circuit which performs
a grasping task, whose architecture and training methods were designed specifically
to result in naturalistic dynamics.

The simulated circuit, from Michaels et al. \cite{michaels.mrnn}, was trained
to resemble the grasping circuits of monkey subjects engaged in a delayed reach-to-grasp task.
Its design draws on a body of literature focused on architectures and training methods for RNNs which
seek to create artificial neural circuits that have activation dynamics similar to natural circuits,
including delayed grasping tasks \cite{susillo.mrnn}. The Michaels circuit consists of a
``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual features
intended to capture the view the monkeys had during the task, i.e. 3D renderings of the same
objects which the monkeys grasped. The outputs are muscle length velocities for the should,
arm, and hand of the monkey during the trial. The natural velocities were captured with a
motion capturing glove, and the artificial neural network was trained to recapitulate those
grasping motions. Data and trained models from this work were supplied to us by the lead
author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.

@@

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
refers to the fact that the activity of the module receiving the visual inputs resembles
the AIP activity of the monkey subject from the same trials. Likewise, the second and
third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
The emergent network dynamics show a number of other correspondences to the monkeys' natural
brain activity as well, as detailed in that paper.

Importantly, the simulated circuit's activity shows a relatively clear separation of the
object shapes. That is - the visual information input to the network that differentiates
one trial from another is leveraged by the circuit to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and
location. The visual information forward-propagates through the network, through successive
processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
for the object represented by the input visual features. It follows, then, that the classes
are separable in some way by observing the distinct ways that the mRNN is activated by the
classes' corresponding visual inputs. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

\subsubsection{Simulated lesions cause real world failure modes}
Simulating a brain lesion in terms of this artificial network results in error
modes that resemble a natural lesion of certain regions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. Conversely, if we simulate a
``stroke'' of a portion of the output module of the Michaels mRNN, roughly
corresponding to M1, we see a more wholesale loss of movement, affecting even the
ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.

\begin{figure}
	\centering
	\begin{subfigure}[c]{0.65\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
	\end{subfigure}
	\hfill
	\begin{subfigure}[c]{0.32\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
	    \caption{Proportion of loss attributable to hand muscles}
	\end{subfigure}
	\hfill
	\caption{Lesion designs which prevent forward propagation of object shape information
	         differentially impact hand pose. Loss here is a mean squared error (MSE) loss,
		 measured relative to the circuit's trajectories prior to the lesion.}
	\label{fig:lesion}
\end{figure}

The co-processor's task in our simulation will be to identify the appropriate information,
read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
to effectively bridge across the lesion, forward propatating the object shape information,
indirectly, through the use of stimulation.

In our experiments below, we demonstrate our co-processor design on two types of
simulated lesions:
\begin{itemize}
	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
	      information between the ``F5'' and ``M1'' modules, effectively representing a
	      severing of the connections between the two.
	\item \textbf{Output-based:} we force the output of some proportion of ``M1'' neurons
	      to zero, effectively removing them from the network.
\end{itemize}

TODO: show example trajectories for both

\subsubsection{Simulated network exhibits long running dynamics}
Just as the co-processor must adapt to

\subsubsection{Sensor model}
asdf

\subsubsection{Stimulation model}
asdf
